---
title: "Assignment B1"
author: "Cynthia"
date: "2025-10-24"
output: html_document
---

## Exercise 1: Make a Function (25 points)

In this exercise, you’ll be making a function and fortifying it. The function need not be complicated. The function need not be “serious”, but shouldn’t be nonsense.

Function Ideas

-   Did you repeat any code for a data analysis in STAT 545A? If so, consider making a function for this action.

-   Consider bundling a specific group_by() %\>% summarise() workflow.

-   Write a wrapper around an existing function.

    -   – For example, perhaps accepting a narrower range of inputs (like not allowing logical vectors), or providing a different output.

    -   – A specific example: the rqdist() function is a wrapper around quantreg::rq(), narrowing its functionality.

    -   – It’s usually better to narrow a function’s focus than to broaden, so that a function doesn’t end up doing too much.

-   Make a special plot that you’d want to repeat when exploring your data.

-   . . .

Guidelines

-   Your function should not rely on anything from your working environment.

-   Your function should not rely on “magic numbers”

    -   – pre-selected numbers (or options) that appear inside the function that can’t be accessed by a user of the function.

    -   – For example, maybe quantile(x, type = 1, ...) appears in your function. The choice of 1 is arbitrary – unless you’re making a function like quantile_type_1().

-   Your function should be reasonably flexible on its inputs. Here is an example of a function that is not reasonably flexible: a function called count_missing_values_by_group() that only works when the data frame has columns that are all of type fct.

-   The output is consistent – for example, always gives a list. An example of inconsistent output: sapply(1:3, seq_len) gives a list, and sapply(1:3, sqrt) gives an (atomic) vector.

Your function includes appropriate arguments. (Do you handle NA’s appropriately? Are you using the ellipsis properly? etc.)

```{r}
#Load appropriate packages
library(dplyr)
library(datateachr)
library(testthat)
```

```{r}
#' Summarise a numeric column by group
#'
#' This function takes a data frame and summarises a numeric column
#' (ex. income, score) by a chosen grouping variable (ex. region, category). 
#' It returns the mean and median of the numeric variable for each group.
#'
#' @param data A data frame containing the variables of interest.
#' @param group_col The column used to group observations (e.g., decade or region).
#' @param value_col The numeric column to summarise (e.g. income, score).
#'
#' @return A tibble containing one row per group with two summary statistics:
#' the mean and median of the selected numeric column.

summarise_by_group <- function(data, group_col, value_col) {
  data %>%
    group_by({{ group_col }}) %>%
    summarise(
      mean = mean({{ value_col }}, na.rm = TRUE),
      median = median({{ value_col }}, na.rm = TRUE)
    )
}
```

## Exercise 2: Document your Function (20 points)

In the same code chunk where you made your function, document the function using roxygen2 tags. Be sure to include:

1.  Title.
2.  Function description: In 1-2 brief sentences, describe what the function does.
3.  Document each argument with the @param tag, making sure to justify why you named the parameter as you did. • (Justification for naming is not often needed, but we want to hear your reasoning.)
4.  What the function returns, using the @return tag.

## Exercise 3: Include examples (15 points)

Demonstrate the usage of your function with a few (2) examples. Use one or more new code chunks, describing what you’re doing. Note: If you want to deliberately show an error, you can use error = TRUE in your code chunk option.

### Example 1:

```{r}
head(flow_sample)
summarise_by_group(flow_sample, year, flow)
```

### Example 2:

```{r}
head(mtcars)
summarise_by_group(mtcars, gear, wt)
```

## Exercise 4: Test the Function (25 points)

Running examples is a good way of checking by-eye whether your function is working as expected. But, having a formal “yes or no” check is useful when you move on to other parts of your analysis. Write formal tests for your function. You should use at least three non-redundant uses of an expect\_() function from the testthat package, and they should be contained in a test_that() function (or more than one). They should all pass.

Example of non-redundant inputs:

-   Vector with no NA’s

-   Vector that has NA’s

-   Vector of a different type (if relevant)

-   Vector of length 0, like numeric(0).

Example of redundant inputs:

-   Providing a different number (unless one of these numbers have some significance, like an extreme point – just tell us if that’s the case)

### Test #1:

```{r}
any(is.na(mtcars)) #mtcars has no NA values

# Test 1: Works with complete data
test_that("Function works with vector without NA's", {
  expect_
```

### Test #2:

```{r}
# Test 2: Handles NA values properly
test_that("Function correctly handles NA values", {
  test_data <- tibble(
    group = c("A", "A", "B", "B"),
    value = c(1, NA, 3, 4)
  )
  result <- summarise_by_group(test_data, "group", "value")

  expect_false(any(is.na(result$mean)))              # Means should not be NA
  expect_equal(result$mean[result$group == "A"], 1)  # Only one valid value
})
```

### Test #3:

```{r}
# Test 3: Works with empty data frame
test_that("Function works with empty data", {
  test_data <- tibble(group = character(), value = numeric())
  result <- summarise_by_group(test_data, "group", "value")

  expect_equal(nrow(result), 0)                      # Should return empty tibble
  expect_s3_class(result, "tbl_df")                  # Should still be a tibble
})
```
