---
title: "Assignment B1"
author: "Cynthia"
date: "2025-10-24"
output: github_document
---

## Exercise 1: Make a Function (25 points)

In this exercise, you’ll be making a function and fortifying it.

```{r}
#Load appropriate packages
library(roxygen2)
library(tidyverse)
library(datateachr)
library(testthat)
```

```{r}
#' @title Summarise a numeric column by group
#'
#' @description This function takes a data frame and summarises a numeric column
#' (ex. income, score) by a chosen grouping variable (ex. region, category). 
#' It returns the mean and median of the numeric variable for each group.
#'
#' @param data A data frame containing the variables of interest.
#' @param group_col The column used to group observations (e.g., decade or region). Provides our 'groups'.
#' @param value_col The numeric column to summarise (e.g. income, score). Provides our 'values'.
#'
#' @return A tibble containing one row per group with two summary statistics:
#' the mean and median of the selected numeric column. This function will automatically remove the NAs from your data and thus your resulting tibble will not contain NAs. 


summarise_by_group <- function(data, group_col, value_col) {
  
  # Shows error if group_col or value_col is in string format (eg. "group_col")
  if (any(is.character(substitute(group_col)), is.character(substitute(value_col)))) { #substitute allows us to directly view what the user typed as input, this will stop the function if either group_col or value_col is a string
    stop("Please provide column names without quotes")
  }
  
  # Show error if value_col is not numeric
  if(!is.numeric(pull(data, {{value_col}}))){
    stop("The value_col column must be numeric")
  }
  
  data %>%
    group_by({{ group_col }}) %>%
    summarise(
      mean = mean({{ value_col }}, na.rm = TRUE),
      median = median({{ value_col }}, na.rm = TRUE),
      .groups = "drop" #drops grouping to avoid issues in future use
    )
}

```

## Exercise 2: Document your Function (20 points)

In the same code chunk where you made your function, document the function using roxygen2 tags. Be sure to include:

1.  Title.
2.  Function description: In 1-2 brief sentences, describe what the function does.
3.  Document each argument with the @param tag, making sure to justify why you named the parameter as you did. • (Justification for naming is not often needed, but we want to hear your reasoning.)
4.  What the function returns, using the @return tag.

## Exercise 3: Include examples (15 points)

Demonstrate the usage of your function with a few examples. Use one or more new code chunks, describing what you’re doing.

### Example 1: Applying the function to the flow_sample dataset

Here we are using our function to group by month, and then obtain the mean and median flow values for each month.

```{r}
head(flow_sample)
summarise_by_group(flow_sample, month, flow)
```

### Example 2: Applying the function to the mtcars dataset

Here we are grouping by gear number, and then calculating the mean and median weight for each group.

```{r}
head(mtcars)
summarise_by_group(mtcars, gear, wt)
```

## Exercise 4: Test the Function (25 points)

Running examples is a good way of checking by-eye whether your function is working as expected. But, having a formal “yes or no” check is useful when you move on to other parts of your analysis. Write formal tests for your function. You should use at least three non-redundant uses of an expect\_() function from the testthat package, and they should be contained in a test_that() function (or more than one). They should all pass.

### Creating tibbles for testing

```{r}

#Tibble for Test 1
my_data1 <- tibble(
  group = c("A", "A", "A", "B", "B", "B", "C", "C", "C"),
  category = c("X", "Y", "Z", "X", "Z", "Y", "Y", "Z", "X"),
  values = c("9", "0", "0", 12, 89, 56, 34, 78, 90)
)

my_data1

#Tibble for Test 2
my_data2 <- tibble(
  group = c("A", "A", "A", "B", "B", "B", "C", "C", "C"),
  category = c("X", "Y", "Z", "X", "Z", "Y", "Y", "Z", "X"),
  values = c(NA, NA, NA, 12, 89, 56, 34, 78, 90)
)

my_data2

#Tibble for Test 3
my_data3 <- tibble(
  group = c("A", "A", "A", "B", "B", "B", "C", "C", "C"),
  category = c("X", "Y", "Z", "X", "Z", "Y", "Y", "Z", "X"),
  values = c(9, 0, 0, 12, 89, 56, 34, 78, 90)
)

my_data3
```

### Test #1: Does not take non-numeric values as input for value_col

```{r}

test_that("Function does not accept non-numeric values when summarising", {
  #We would expect an error as our input is non-numeric (it's a string)
  expect_error(summarise_by_group(my_data1, group, values))
 
})

```

### Test #2: Handles NA values correctly

```{r}
             
test_that("Function correctly incorporates NA values when summarising", {
  result <- summarise_by_group(my_data2, group, values)
  
  #Since all values for group A are NA, we should get NA for mean and median
  expect_true(is.na(result %>% filter(group == "A") %>% pull(mean))) #this allows us to filter for only group A's mean value
  expect_true(is.na(result %>% filter(group == "A") %>% pull(median))) #this allows us to filter for only group A's median value
 
})
```

### Test #3: Does not allow user to input strings as column names

```{r}

test_that("Function correctly does not allow user to input strings when referencing columns", {
  
  #We would expect an error as we are not inputting the correct format for group_col
  expect_error(summarise_by_group(my_data3, "group", values))
})
```
