---
title: "Assignment B1"
author: "Cynthia"
date: "2025-10-24"
output: html_document
---

## Exercise 1: Make a Function (25 points)

In this exercise, you’ll be making a function and fortifying it. The function need not be complicated. The function need not be “serious”, but shouldn’t be nonsense.

Function Ideas

-   Did you repeat any code for a data analysis in STAT 545A? If so, consider making a function for this action.

-   Consider bundling a specific group_by() %\>% summarise() workflow.

-   Write a wrapper around an existing function.

    -   – For example, perhaps accepting a narrower range of inputs (like not allowing logical vectors), or providing a different output.

    -   – A specific example: the rqdist() function is a wrapper around quantreg::rq(), narrowing its functionality.

    -   – It’s usually better to narrow a function’s focus than to broaden, so that a function doesn’t end up doing too much.

-   Make a special plot that you’d want to repeat when exploring your data.

-   . . .

Guidelines

-   Your function should not rely on anything from your working environment.

-   Your function should not rely on “magic numbers”

    -   – pre-selected numbers (or options) that appear inside the function that can’t be accessed by a user of the function.

    -   – For example, maybe quantile(x, type = 1, ...) appears in your function. The choice of 1 is arbitrary – unless you’re making a function like quantile_type_1().

-   Your function should be reasonably flexible on its inputs. Here is an example of a function that is not reasonably flexible: a function called count_missing_values_by_group() that only works when the data frame has columns that are all of type fct.

-   The output is consistent – for example, always gives a list. An example of inconsistent output: sapply(1:3, seq_len) gives a list, and sapply(1:3, sqrt) gives an (atomic) vector.

Your function includes appropriate arguments. (Do you handle NA’s appropriately? Are you using the ellipsis properly? etc.)

```{r}
#Load appropriate packages
library(dplyr)
library(datateachr)
library(testthat)
```

```{r}
#' @title Summarise a numeric column by group
#'
#' @description This function takes a data frame and summarises a numeric column
#' (ex. income, score) by a chosen grouping variable (ex. region, category). 
#' It returns the mean and median of the numeric variable for each group.
#'
#' @param data A data frame containing the variables of interest.
#' @param group_col The column used to group observations (e.g., decade or region).
#' @param value_col The numeric column to summarise (e.g. income, score).
#'
#' @return A tibble containing one row per group with two summary statistics:
#' the mean and median of the selected numeric column. This function will automatically remove the NAs from your data and thus your resulting tibble will not contain NAs

summarise_by_group <- function(data, group_col, value_col) {
  data %>%
    group_by({{ group_col }}) %>%
    summarise(
      mean = mean({{ value_col }}, na.rm = TRUE),
      median = median({{ value_col }}, na.rm = TRUE)
    )
}
```

## Exercise 2: Document your Function (20 points)

In the same code chunk where you made your function, document the function using roxygen2 tags. Be sure to include:

1.  Title.
2.  Function description: In 1-2 brief sentences, describe what the function does.
3.  Document each argument with the @param tag, making sure to justify why you named the parameter as you did. • (Justification for naming is not often needed, but we want to hear your reasoning.)
4.  What the function returns, using the @return tag.

## Exercise 3: Include examples (15 points)

Demonstrate the usage of your function with a few (2) examples. Use one or more new code chunks, describing what you’re doing. Note: If you want to deliberately show an error, you can use error = TRUE in your code chunk option.

### Example 1:

```{r}
head(flow_sample)
summarise_by_group(flow_sample, year, flow)
```

### Example 2:

```{r}
head(mtcars)
summarise_by_group(mtcars, gear, wt)
```

## Exercise 4: Test the Function (25 points)

Running examples is a good way of checking by-eye whether your function is working as expected. But, having a formal “yes or no” check is useful when you move on to other parts of your analysis. Write formal tests for your function. You should use at least three non-redundant uses of an expect\_() function from the testthat package, and they should be contained in a test_that() function (or more than one). They should all pass.

Example of non-redundant inputs:

-   Vector with no NA’s

-   Vector that has NA’s

-   Vector of a different type (if relevant)

-   Vector of length 0, like numeric(0).

Example of redundant inputs:

-   Providing a different number (unless one of these numbers have some significance, like an extreme point – just tell us if that’s the case)

### Creating tibbles for testing

```{r}

#Tibble for Test 1
my_data1 <- tibble(
  group = c("A", "A", "A", "B", "B", "B", "C", "C", "C"),
  category = c("X", "Y", "Z", "X", "Z", "Y", "Y", "Z", "X"),
  values = c(9, 0, 0, 12, 89, 56, 34, 78, 90)
)

my_data1

#Tibble for Test 2
my_data2 <- tibble(
  group = c("A", "A", "A", "B", "B", "B", "C", "C", "C"),
  category = c("X", "Y", "Z", "X", "Z", "Y", "Y", "Z", "X"),
  values = c(NA, NA, NA, 12, 89, 56, 34, 78, 90)
)

my_data2

#Tibble for Test 3
my_data3 <- tibble(
  group = c("A", "A", "A", "B", "B", "B", "C", "C", "C"),
  category = c("X", "Y", "Z", "X", "Z", "Y", "Y", "Z", "X"),
  values = c(2, -4, -1, 12, 89, 56, 34, 78, 90)
)

my_data3

```

### Test #1: Handles values of zero correctly

```{r}

test_that("Function correctly incorporates 0 values when summarising", {
  result <- summarise_by_group(my_data1, group, values)
  
  #Manually calculating the value, we should get 3 for mean and 0 for median
  
  #Expect manually calculated value in the result
  expect_equal(result %>% filter(group == "A") %>% pull(mean), 3) #this allows us to filter for only group A's mean value
  expect_equal(result %>% filter(group == "A") %>% pull(median), 0) #this allows us to filter for only group A's meedian value
 
})

```

### Test #2: Handles NA values correctly

```{r}
             
test_that("Function correctly incorporates NA values when summarising", {
  result <- summarise_by_group(my_data2, group, values)
  
  #Since all values for group A are NA, we should get NA for mean and median
  
  #Expect manually calculated value in the result
  expect_true(is.na(result %>% filter(group == "A") %>% pull(mean))) #this allows us to filter for only group A's mean value
  expect_true(is.na(result %>% filter(group == "A") %>% pull(median))) #this allows us to filter for only group A's meedian value
 
})
```

### Test #3: Handles negative values correctly

```{r}
test_that("Function correctly incorporates negative values when summarising", {
  result <- summarise_by_group(my_data3, group, values)
  
  #Manually calculating the value, we should get -1 for mean and -1 for median
  
  #Expect manually calculated value in the result
  expect_equal(result %>% filter(group == "A") %>% pull(mean), -1)
  expect_equal(result %>% filter(group == "A") %>% pull(median), -1)
 
})
```
